---
title: "Bioinformatics Class 9"
author: "Matt DuPrie"
date: "May 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

# LOAD DATA

url <- "https://bioboot.github.io/bggn213_S18/class-material/WisconsinCancer.csv"

# use read.csv to import comma separated file
wisc.df <- read.csv(url)


# How many diagnosis are cancer vs. non-cancer
# table(wisc.df$diagnosis)



```


```{r}
# Make a new matrix that gets rid of first two columns, which contain patient ID and diagnosis factor
# Use - in following code to specify to delete certain columns
## Must also delete final column 33 due to non-numerical input

wisc.data <- as.matrix(wisc.df[,-c(1:2,33)])
rownames(wisc.data) <- wisc.df$id


x <- c("barry", "chris", "mary", "pete")
```


Finally, setup a separate new vector called diagnosis to be 1 if a diagnosis is malignant ("M") and 0 otherwise. Note that R coerces TRUE to 1 and FALSE to 0
```{r}

diagnosis <- as.numeric(wisc.df$diagnosis=="M")
sum(diagnosis)
```


Begin analyzing data
```{r}
#Q1. How many observations are in this dataset?
# use function nrows, gives total # of rows
# can use dim
nrow(wisc.data)

#Q2. How many variables/features in the data are suffixed with _mean?
length(grep("mean", colnames(wisc.data), value = TRUE ))


#Q3. How many of the observations have a malignant diagnosis?

diagnosis <- as.numeric(wisc.df$diagnosis=="M")
sum(diagnosis)


```


Now try PCA, must determine if scaling necessary

```{r}
# Check column means and standard deviations
colMeans(wisc.data)
plot( colMeans(wisc.data), type = "o")
# Yes, you should scale, large variance in means


#this determines sd for data
apply(wisc.data,2,sd)
```



Perform PCA

```{r}
wisc.pr <- prcomp(wisc.data, scale = TRUE)

summary(wisc.pr)
```


PCA Results

```{r}
plot(wisc.pr$x[,1], wisc.pr$x[,2], col= diagnosis+1)
```


Scree-plot: vairance explained.  Choose fewest PCs that explain the most variance.  Choose first 2,3.

```{r}
#R won't give variance directly, must create new vector that has variance.  Variance is stdev^2.

pve <- (wisc.pr$sdev^2)/sum(wisc.pr$sdev^2)

plot(pve, xlab = "Principal Component",
     ylab = "Proportion of Variance Explained",
     ylim = c(0,1), type = "o")
```


Bar plot of data
```{r}
# Alternative scree plot of the same data, note data driven y-axis
barplot(pve, ylab = "Precent of Variance Explained",
     names.arg=paste0("PC",1:length(pve)), las=2, axes = FALSE)
axis(2, at=pve, labels=round(pve,2)*100 )
```








SECTION 3 HIERARCHICAL CLUSTERING OF CASE DATA


```{r}

data.scaled <- scale(wisc.data)

data.dist <- dist(data.scaled)

wisc.hclust <- hclust(data.dist, method = "complete")

plot(wisc.hclust)
abline(h=20, col = "red", lwd = 3)
```



```{r}


wisc.hclust.clusters <- cutree(wisc.hclust, k = 4)
#plot(wisc.hclust.clusters)
table(wisc.hclust.clusters, diagnosis)
```



K-means clustering

```{r}
wisc.km <- kmeans(wisc.data, centers= 2, nstart= 20)

table(wisc.km$cluster, diagnosis)


```


Compare the 2 methods
```{r}
table(wisc.km$cluster,wisc.hclust.clusters)
```

Clustering on PCA results

```{r}
# Call clustering method on distance matrix on first 7 PCs
wisc.pr.hclust <- hclust(dist(wisc.pr$x[, 1:2]), method="ward.D2")

wisc.pr.hclust.clusters <- cutree(wisc.pr.hclust, k=2)

table(wisc.pr.hclust.clusters, diagnosis)
plot(wisc.pr$x[,1:2], col=wisc.pr.hclust.clusters)
```


```{r}

```

BONUS

```{r}
## Predicting malignancy of new samples

url <- "https://tinyurl.com/new-samples-CSV"
new <- read.csv(url)
npc <- predict(wisc.pr, newdata=new)

plot(wisc.pr$x[,1:2], col=wisc.pr.hclust.clusters+1)
points(npc[,1], npc[,2],col= "blue", pch=16, cex=3)
```









