---
title: "Bioinformatics Class 7"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions again

We can source any file of R code with the 'source'() function

```{r}
source("http://tinyurl.com/rescale-R")
```

let's make sure things are here
```{r}
ls()

```

Check our rescale function is working

```{r, eval=FALSE}
rescale(c(1:10, "string"))
```

Let's check if 'rescale2()' does any better

```{r, eval=FALSE}
rescale2(c(1:10, "string"))
```



## Function for finding  missing values in two datasets

Write a 'both_na()' function to do this...
```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)
is.na(x)
```

Function determines which positions are NA
```{r}
which( is.na(x) )

```
```{r}
is.na(x)
is.na(y)
```

Now want to write function that finds at what position both series x and y are NA, sum will give total amount of positions where this is true. 
```{r}
sum( is.na(x) & is.na(y))
```

My first function can start from this snippet :-)

```{r}
both_na2 <- function(x, y) {
  ##Check for NA elements in both input vectors and don't allow
    if(length(x) != length(y)) {
   stop("Input x and y should be the same length")
    }
  sum( is.na(x) & is.na(y) )
}
```

```{r}
x <-  c(NA, NA, NA)
y1 <- c( 1, NA, NA)
y2 <- c( 1, NA, NA, NA)


both_na2(x, y1, y2)
```


```{r}
function(x,y) {
  ##Check for NA elements in both input vectors and don't allow
    if(length(x) != length(y)) {
   stop("Input x and y should be the same length")
    }
  sum( is.na(x) & is.na(y))
}
```
```{r}
both_na2(x, y2)
```


```{r}
both_na3 <- function(x, y) {
  if(length(x) != length(y)) {
    stop("Input x and y should be vectors of the same length")
  }
  na.in.both <- ( is.na(x) & is.na(y) )
  na.number  <- sum(na.in.both)
  na.which   <- which(na.in.both)
  message("Found ", na.number, " NA's at position(s):", 
          paste(na.which, collapse=", ") ) 
  return( list(number=na.number, which=na.which) )
}
```

```{r}
x <- c( 1, 2, NA, 3, NA)
y <- c(NA, 3, NA, 3,  4)

both_na3(x,y)
```


```{r}
ans$which
```


## And a last function that is actually useful

```{r}
x <- df1$IDs
y <- df2$IDs

x
y
```

We can try the 'intersect()' function and the '%in%' functions

```{r}
intersect(x,y)
x %in% y
```

We can use logical output in '%in%' to get at our data

```{r}
x[x %in% y]
y[y %in% x]
```


Let's put these together as columns of a matrix


```{r}
cbind( x[ x %in% y ], y[ y %in% x ] )
```


```{r}
cbind(c("Hello", "Help"), c("Please", "Me"))
```


Make new function
```{r}
gene_intersect <- function(x, y) { 
   cbind( x[ x %in% y ], y[ y %in% x ] )
}

gene_intersect(x,y)
```

Let's try with data.frame input rather than vectors

```{r}
gene_intersect2 <- function(df1, df2) { 
   cbind( df1[ df1$IDs %in% df2$IDs, ], 
          df2[ df2$IDs %in% df1$IDs, "exp"] )
}
```

```{r}

gene_intersect2(df1, df2)
```




