---
title: "Lecture 11"
author: "Matt DuPrie"
date: "May 9, 2018"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB statistics

Import our PDB statistics CSV file and calculate percent structures by experimental method.


```{r}
p <- read.csv("Data Export Summary.csv", row.names = 1)
```



```{r}

percent <- (p$Total / sum(p$Total)) * 100
names(percent) <- row.names(p)
percent

```


Using bio3D

Load the bio3D package....
type install.packages(bio3D) if you get error message
to list all functions:
bio3d()
to get help for all functions:
help(package=bio3d)

```{r}
library(bio3d)
```

Read our HIV-Pr structure


```{r}
pdb <-read.pdb("1hsg")
pdb
```


Q8. Use the Bio3D write.pdb() function to write out a new C-alpha atom only PDB file for viewing in VMD. 

```{r}
# Use select atom function to select only c-alpha
ca.1hsg <-atom.select(pdb, "calpha")
head(ca.1hsg)

#write.pdb(pdb, file = new_pdb)



```

W/ instructor help

```{r}
inds.ligand <-  atom.select(pdb, "ligand")
inds.protein <- atom.select(pdb, "protein")
inds.protein


```



Check we have what we want

```{r}
pdb$atom[inds.ligand$atom,]
```


```{r}
head(pdb$atom[inds.protein$atom,])
```

```{r}

pdb.ligand <- trim.pdb(pdb, inds= inds.ligand)
pdb.protein <- trim.pdb(pdb, inds=inds.protein)

```

```{r}
write.pdb(pdb.ligand, file="1hsg_ligand.pdb")
write.pdb(pdb.protein, file="1hsg_protein.pdb")
```


```{r}
ids <-c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")
files <-get.pdb(ids, split =TRUE)
```


```{r}
# Extract and align the chains we are interested in
pdbs <-pdbaln(files, fit =TRUE)

# Print to screen a summary of the 'pdbs' object
pdbs
```



```{r}
# Access the first 5 rows, and 8 columns
pdbs$ali[1:5, 1:8]

# Associated residues numbers
pdbs$resno[1:5, 1:8]

# Inspect alignment for gaps
gaps <-gap.inspect(pdbs$ali)
# Indices of non-gap columns
head(gaps$f.inds)
# Access non-gap columns
pdbs$ali[, gaps$f.inds]
```


```{r}
# Calculate sequence identity
seqidentity(pdbs)

# Calculate RMSD
rmsd(pdbs)
```

```{r}
# Calculate RMSD
rd <-rmsd(pdbs)

# Clustering 
hc <-hclust(as.dist(rd))
grps <-cutree(hc, k=3)

# Plot results as dendrogram
hclustplot(hc, k=3)
```


```{r}
# Perform PCA
pc <-pca(pdbs)
plot(pc, col=grps)
```




## 5.1 Search and retrieve Adenylate kinase structures

```{r}
aa <-get.seq("1ake_A")
aa

```

```{r}
# Blast or hmmer search 
b <-blast.pdb(aa)
b
```


```{r}
# Plot a summary of search results
hits <-plot(b)

```



```{r}
head(hits$pdb.id)
```




```{r}
hits
```



```{r}
# Fetch PDBs

files <-get.pdb(hits$pdb.id, path ="pdbs", split =TRUE, gzip =TRUE)

# Align structures
pdbs <-pdbaln(files)

# Vector containing PDB codes
ids <-basename.pdb(pdbs$id)

# Draw schematic alignment 
plot(pdbs, labels=ids)
```


```{r}
#5.2  Sequence conservation analysis
# Calculate sequence conservation
cons <-conserv(pdbs, method="entropy22")

# SSE annotations
sse <-pdbs2sse(pdbs, ind=1, rm.gaps=FALSE)

# Plot conservation per residue
plotb3(cons, sse=sse, ylab="Sequence entropy")
```



```{r}
# find invariant core
core <-core.find(pdbs)

# superimpose all structures to core
pdbs$xyz =pdbfit(pdbs, core)

# Perform PCA
pc.xray <-pca(pdbs)

# Calculate RMSD
rd <-rmsd(pdbs)

# Structure-based clustering
hc.rd <-hclust(dist(rd))grps.rd <-cutree(hc.rd, k=3)plot(pc.xray, 1:2, col="grey50", bg=grps.rd, pch=21, cex=1)
```


```{r}
# Left-click on a point to label and right-click to end
identify(pc.xray$z[,1:2], labels=basename.pdb(pdbs$id))

# Visualize first principal component
mktrj(pc.xray, pc=1, file="pc_1.pdb")
```


```{r}
#5.5  Plotting results with ggplot2
library(ggplot2)
library(ggrepel)
df <-data.frame(x=pc.xray$z[,1], y=pc.xray$z[,2])
col <-as.factor(grps.rd)

p <-ggplot(df, aes(x, y)) +geom_point(aes(col=col), size=2) +xlab("PC1") +ylab("PC2") +scale_color_discrete(name="Clusters") +geom_text_repel(aes(label=ids))
```







